---
title: "Remotely sensed forest understory density and nest predator occurrence interact to predict suitable breeding habitat and the occurrence of a resident boreal bird species - Part 1"
author: "Julian Klein"
date: "14 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/RDefaultDirectory/RProjects/Lavskrika/lavskrika_rep_success")
```

# Part 1: LiDAR data and distance to settlement calculation

The is based on rasters data extracted from .las files with FUSION. As raw LiDAR data cannot be redistributed part 1 cannot be run without first downloading LiDAR data from www.lantmateriet.se. The download is not free.

Also the shape files showing the study area and the forestry interventions cannot be made available online. The shapes are nevertheless shown viually in this document. The data set with the exact position of the nests and the human settlement is confidiential as well. Therefore no data used for this analysis is available online. The output of this script is available online and used for part 2 & 3.

Some terms in this analysis:

ALS = LiDAR data, vd = vegetation density, dts = distance to closest human setllement

# Start:

1. Load all packages:

```{r, message=F, warning=F}
library(data.table)
library(raster)
library(dplyr)
```

2. Define all functions: This function extracts the mean of all data from ALS_by_year within i metres.

```{r}
extract_around_nest <- function(x) {
  
  ## We want to extract all data in ALS_by_year within i metres around each nest
  all <- extract(ALS_by_year[[paste0("y_", x$year)]], x[, c("X", "Y")], buffer = i)
  all <- as.data.frame(all)
  
  ## dts is extracted only at the nest since it is the same for all 
  dts <- extract(ALS_by_year[[paste0("y_", x$year)]]$dts, x[, c("X", "Y")])
  
  ## We need the ratio of no data within i metres around the nest
  nd_ratio <- sum(na.omit(all[, 1] == -9999))/sum(!is.na(all[, 1]))
  
  if(nd_ratio < nd_thresh) {
    ## If less than nd_thresh the mean for all ALS metrics is calculated
    all[all == -9999] <-  NA
    out <- c(colMeans(all[, - length(all)], na.rm = TRUE), "dts" = dts)
  } else {
    ## If than nd_thresh NA is returned, except for dts
    all[] <- NA
    out <- c(colMeans(all[, - length(all)]), "dts" = dts)
  }
  
  ## The data for nest x is returned
  return(cbind(x, t(out)))
  
}
```

3. Load all data:

```{r, message=F, warning=F}
## Text files
nest_pos <- read.csv("data/nest_positions.csv")
settl <- read.csv("data/settlement_positions.csv")

## Raster files
ALS <- merge(stack(c("data/unmanaged_ElevP95.asc",
                     "data/unmanaged_5_Perc_above.asc",
                     "data/unmanaged_0.5_Perc_above.asc")),
             stack(c("data/managed_ElevP95.asc",
                     "data/managed_5_Perc_above.asc",
                     "data/managed_0.5_Perc_above.asc")))
names(ALS) <- c("height", "vd_5to", "vd_0to")

## Shape files
study_area <- shapefile("data/120ha_buffer_study.shp")
forestry <- shapefile("data/forestry.shp")

head(nest_pos)
```

```{r}
plot(ALS)
plot(study_area, add = TRUE) 
```

```{r}
plot(forestry)
```

4. Process ALS data and adjust it to forestry interventions:

```{r, message=F, warning=F}
## Calculate absolute percentage returns between 0.5 and 5m
ALS$vd_0to5 <- ALS$vd_0to - ALS$vd_5to

## Reduce ALS data to needed layers
ALS <- ALS[[c(1, 4)]]

## Reduce ALS data set to study area
ALS <- mask(ALS, study_area)

## Add forest area layer: All pixels with forest height below 2 m or NA(water)
ALS$area <- !(ALS[["height"]][] < 2 | is.na(ALS[["height"]][]))

## Create a ALS data sample from all clear cuts in the study area. We use this later on in the script
sample_cc <- mask(ALS, forestry[forestry@data$GRIDCODE %in% as.character(1997:2010) & forestry@data$action == "cut", ])

## Duplicate ALS data from collection year in 2010 onwards for all years we have nest data for

## Create list for yearly ALS data for which nest data exists and store ALS data in it
ALS_by_year <- vector("list", length(c(1998:2004, 2011:2013)))
ALS_by_year[1:length(ALS_by_year)] <- ALS ## Ignore warning!
names(ALS_by_year) <- paste0("y_", c(1998:2004, 2011:2013))

## In the following loop we replace pixels that were cut after collection year 2010 with random samples from clear cuts from 1997 to 2009.

T1 <- NULL

for(i in 2011:2013) {
  
  ## Store ALS data used in this loop in T1 to make understanding easier
  T1 <- ALS_by_year[[paste0("y_", i)]]
  
  ## Which forestry shapes were cut between collection year 2010 and year i
  B1 <- forestry@data$GRIDCODE %in% as.character(2010:i) & forestry@data$action == "cut"
  
  ## Replace all pixels not NA which were cut after collection year 2010 with 
  ## random samples from clear cuts from 1997 to 2009.
  if(sum(B1) > 0) {
    B2 <- !is.na(mask(T1, forestry[B1, ])[[1]][])
    T1[B2] <- sampleRandom(sample_cc, sum(B2))
  } 
  
  ## Update forest are layer with new height data on modelled clear cuts
  T1$area <- !(T1[["height"]][] < 2 | is.na(T1[["height"]][]))
  
  ## Store T1 in respective place in ALS_by_year list
  ALS_by_year[[paste0("y_", i)]] <- T1
  
  T1 <- NULL
  
} 
```

5. Add no data (-9999) to ALS layers. Replace pixels with -9999 where no data exists. These pixels are forests older than a clear cut or thinnings that occured before data acquisition in 2010 as well as forests after thinnings that happened after 2010. No data pixels bear the value -9999 because NA is used for water.  

```{r}
## layer names needed for after loop
names <- names(ALS_by_year[[1]])

for(i in unique(nest_pos$year)) {
  
  if(i < 2010) {
    ## All cc and thinnings after year i until 2010 are no data
    ## i+1 because it is the interventions in the years after i that matter
    B3 <- forestry@data$GRIDCODE %in% as.character((i+1):2010)
  } else {
    ## All thinnings after 2010 become no data
    B3 <- forestry@data$GRIDCODE %in% as.character(2010:i) & forestry@data$action == "thinned"
  }  
  
  if(sum(B3) > 0) {
    for(j in 1:length(ALS[1])) {
      ## All forestry shapes which are TRUE in B3 need to become no data
      ALS_by_year[[paste0("y_", i)]][[j]] <- rasterize(forestry[B3, ], ALS_by_year[[paste0("y_", i)]][[j]], update = TRUE, field = "value")
    }
  }
  
  names(ALS_by_year[[paste0("y_", i)]]) <- names
  
}
```

6. Create and add settlement layer to the ALS_by_year. In one location a settlement was abandoned in 2004. Therefore there are different layers until Spring 2004 and after.

```{r}
## Until 2004, because one settlement disappeard after 2004
dts_u04 <- distanceFromPoints(ALS[[1]], as.matrix(settl[, 1:2]))
names(dts_u04) <- "dts"

dts_a04 <- distanceFromPoints(ALS[[1]], as.matrix(settl[settl$Settlement != "Fika_until_2004", 1:2]))
names(dts_a04) <- "dts"

## Add to ALS_by_year
ALS_by_year[paste0("y_", 1998:2004)] <- lapply(ALS_by_year[paste0("y_", 1998:2004)], FUN = function(x) stack(x, dts_u04))
ALS_by_year[paste0("y_", 2011:2013)] <- lapply(ALS_by_year[paste0("y_", 2011:2013)], FUN = function(x) stack(x, dts_a04))
```

7. Extract ALS data from ALS_by_year for different radiuses around the nest and export.

```{r}
nest_pos <- as.data.table(nest_pos)

## Calculate radiuses for fractions of territory size 6.6E5m2
rad <- c(15, ceiling(sqrt(6.6E5/pi)*seq(0.04, 1, 0.01)))

## Define no data threshhold accepted (now 5% no data)
nd_thresh <- 0.05

## Extract data for all radiuses around the nest

ALS_out <- NULL
T2 <- NULL

for(i in rad) {

  T2 <- nest_pos[, extract_around_nest(.SD), by = 1:nrow(nest_pos)] 
  T2$sample_rad <- i

  ALS_out <- rbind(ALS_out, T2)
  
}

head(ALS_out); str(ALS_out)
```

The data set "ALS_rep_succ.csv" is merged with nest data and used in Part 2. 
The territory names seen in the data presented here is encoded in the data which
is published. Also the nest coordinates are removed.