---
title: "Remotely sensed forest understory density and nest predator occurrence interact to predict suitable breeding habitat and the occurrence of a resident boreal bird species - Part 2"
author: "Julian Klein"
date: "14 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/RDefaultDirectory/RProjects/Lavskrika/lavskrika_rep_success")
```

# Part 2: The whole analysis of breeding succes is perfomred in this script. The landscape prediction comparison is in part 3.

# Start:

1. Load all packages:

```{r, message=F, warning=F}
library(data.table)
library(MuMIn)
library(DHARMa)
library(lme4)
library(car)
library(ggplot2)
```

2. Define all functions:

```{r}
## Specify control values for all models in this script
cont_spec <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))

## Test model assumptions
test_my_model <- function(m.out) {
  
  sim <- simulateResiduals(m.out)
  plot(sim)
  print(testUniformity(sim))
  print(testZeroInflation(sim))
  print(testDispersion(sim))
  print("Variance Inflation Factor:")
  print(vif(m.out))
  
}
```

3. Load all data:

```{r}
## ALS data and nest data comnbined
nest_ALS <- read.csv("data/siberian_jay_data.csv")

head(nest_ALS); str(nest_ALS)
```

4. Make models for different dts categorisation and categorise dts according to result.

```{r, message=F, warning=F}
## Reduce nest_ALS to one sample radius
D <- nest_ALS[nest_ALS$sample_rad == 15, ]

## Make loop through different categorisation distances and store results
r.dts_cat <- NULL
for(i in seq(500, 3500, 50)) {

  ## Categorise dts
  D$dts_cat <- ifelse(D$dts > i, "far", "close")

  m.dts_cat <- glmer(rep_succ ~ dts_cat + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year), 
                     family = binomial, 
                     data = D, 
                     control = cont_spec)
  
  ## Store model output for all categorisation distances
  r.dts_cat <- rbind(r.dts_cat, cbind("AIC" = AIC(m.dts_cat), "dts_cat" = i))

}

head(r.dts_cat)
```

```{r}
## Test model assumptions with DHARMa for chosen dts_cat

## Select dts_cat with lowest Estimate
R <- r.dts_cat[r.dts_cat[,"AIC"] == min(r.dts_cat[,"AIC"]), "dts_cat"]
print(R)
```

```{r, message=F, warning=F}
D$dts_cat <- ifelse(D$dts > R, "far", "close")

m.dts_cat <- glmer(rep_succ ~ dts_cat + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year), 
                   family = binomial, 
                   data = D, 
                   control = cont_spec)

tryCatch(test_my_model(m.dts_cat), error = function(e) {cat("ERROR :", conditionMessage(e), "\n")})
```

5. Make the main model testing ALS on breeding success for 15 m radius around the nest, beacuse most data for that one. Test alternative (linear and quadratic instead of logarithmic) relationships between understory density and breeding success with AIC.

```{r, message=F, warning=F}
## Categorise dts according to the results above in nest_ALS
nest_ALS$dts_cat <- ifelse(nest_ALS$dts > R, "far", "close")

## Exclude NA's in data set for use in models below
DD <- na.omit(nest_ALS)

## Reduce DD to D15
D15 <- DD[DD$sample_rad == 15, ]

## Add logarithmic version of vd_0to5
D15$vd0t5_log <- log(D15$vd_0to5)

## Center all continuous variables to avoid covariate correlations
D15$vd0t5_log_c <- D15$vd0t5_log - mean(D15$vd0t5_log)
D15$vd0t5_c <- D15$vd_0to5 - mean(D15$vd_0to5)

## Test log model with veg density as a logarithmic predictor 

m.vd0t5_log_c_15 <- glmer(rep_succ ~ dts_cat * vd0t5_log_c + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year),
                              family = binomial,
                              data = D15,
                              control = cont_spec)

summary(m.vd0t5_log_c_15)
```

```{r, message=F, warning=F}
r.squaredGLMM(m.vd0t5_log_c_15)
```

```{r, message=F, warning=F}
test_my_model(m.vd0t5_log_c_15)
```

```{r, message=F, warning=F}
## Store data set with prediction for figur1 which is produced below
fig1 <- cbind(predict(m.vd0t5_log_c_15, re.form = NA, se.fit = TRUE, type = "response"), 
              D15[, c("rep_succ", "dts_cat", "vd_0to5")])

## Compare to quadratic, linear and intercep only model

## Only the intercept:
m.int <- glmer(rep_succ ~ (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year),
               family = binomial,
               data = D15,
               control = cont_spec)

## Distance to settlement only:
m.dts_cat_15 <- glmer(rep_succ ~ dts_cat + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year),
                      family = binomial,
                      data = D15,
                      control = cont_spec)

## With veg density only as a linear predictor:
m.vd0t5_c_15_only <- glmer(rep_succ ~ vd0t5_c + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year),
                               family = binomial,
                               data = D15,
                               control = cont_spec)
## Test model assumptions:
tryCatch(test_my_model(m.vd0t5_c_15_only), error = function(e) {cat("ERROR :", conditionMessage(e), "\n")})
```

```{r, message=F, warning=F}
## With veg density as a linear predictor in interaction with dts:
m.vd0t5_c_15 <- glmer(rep_succ ~ dts_cat * vd0t5_c + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year),
                          family = binomial,
                          data = D15,
                          control = cont_spec)
## Test model assumptions:
test_my_model(m.vd0t5_c_15)
```

```{r, message=F, warning=F}
## With veg density as a quadratic predictor:
m.vd0t5_c_poly_15 <- glmer(rep_succ ~ dts_cat * poly(vd0t5_c, 2) + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year),
                               family = binomial,
                               data = D15,
                               control = cont_spec)
## Test model assumptions:
test_my_model(m.vd0t5_c_15)
```

```{r, message=F, warning=F}
## List all models above in a model selection table and export results:
model.sel(m.vd0t5_log_c_15, m.int, m.dts_cat_15, m.vd0t5_c_15_only, m.vd0t5_c_15, m.vd0t5_c_poly_15)
```

6. Check wether the results of the main model are sensitive to the categorisation distance of the distance of the nest to the closest settlement.

```{r, message=F, warning=F}
## Make loop through different categorisation distances and store results:
r.sensitivity <- NULL
for(i in seq(500, 3500, 50)) {
  
  ## Categorise dts
  D15$dts_cat <- ifelse(D15$dts > i, "far", "close")
  
  m.sensitivity <- glmer(rep_succ ~ dts_cat * vd0t5_log_c + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year),
                         family = binomial,  
                         data = D15,
                         control = cont_spec)
  
  ## Store model output for all categorisation distances:
  r.sensitivity <- rbind(r.sensitivity, 
                         cbind(summary(m.sensitivity)$coefficients, 
                               "R2m" = r.squaredGLMM(m.sensitivity)[1, 1], 
                               "AIC" = AIC(m.sensitivity), 
                               "dts_cat" = i))
  
}

head(r.sensitivity)
```

7. Test the logarithmic model with understory density for all radiuses around the nest and extract AIC values. Also calculate the correlation of the mean vd_0to5 of all radiuses with 15 m.

```{r, message=F, warning=F}
## Reduce data set so the same nests are used for all radiuses:

## Which radius has fewest nests?
N1 <- names(sort(table(DD$sample_rad)))[1]

## Which nests are those?
N2 <- DD[DD$sample_rad == N1 & !is.na(DD$vd_0to5), "name"]

## Select only those nests:
DD_all_rad <- DD[DD$name %in% N2, ]

## Add centered log(vd_0to5) by radius:
DD_all_rad <- as.data.table(DD_all_rad)
DD_all_rad[, "vd0t5_log_c" := log(vd_0to5) - mean(log(vd_0to5)), by = "sample_rad"]

## Make a loop through all radiuses, run the log model and store results:

r.all_rad <- NULL
for(i in unique(DD_all_rad$sample_rad)) {

  tryCatch({
    
    ## Calculate correlation of vd_0to5 at all radiuses with 15m:
    r.cor <- cor(DD_all_rad[DD_all_rad$sample_rad == i, "vd_0to5"],
                 DD_all_rad[DD_all_rad$sample_rad == 15, "vd_0to5"])
  
    m.all_rad <- glmer(rep_succ ~ dts_cat * vd0t5_log_c + area + (1|female_ring) + (1|male_ring) + (1|hab_qual) + (1|year),
                       family = binomial,
                       data = DD_all_rad[DD_all_rad$sample_rad == i, ],
                       control = cont_spec)
  
    ## Store model output for all radiuses:
    r.all_rad <- rbind(r.all_rad, 
                       cbind("cor" = r.cor[1],
                             "estimate" = summary(m.all_rad)$coefficients[5, 1],
                             "SE" = summary(m.all_rad)$coefficients[5, 2],
                             "pvalue" = summary(m.all_rad)$coefficients[5, 4],
                             "AIC" = AIC(m.all_rad), 
                             "R2m" = r.squaredGLMM(m.all_rad)[1, 1],
                             "radius" = i))
  
  }, error = function(e) {cat("ERROR :", conditionMessage(e), "\n")})
  
}

## Add deltaAIC to results:
r.all_rad <- as.data.frame(r.all_rad)
r.all_rad$deltaAIC <- r.all_rad$AIC - min(r.all_rad$AIC)

print(r.all_rad)
```

8. Figure 1:

```{r}
## Change dts_cat level names:
fig1$dts_cat <- as.factor(fig1$dts_cat)
levels(fig1$dts_cat) <- c("close to human settlement", "far from human settlement") 

ggplot(fig1, aes(x = vd_0to5, y = rep_succ, fill = dts_cat, color = dts_cat)) +
geom_jitter(size = 2, na.rm = TRUE, width = 0, height = 0.01) +
geom_line(aes(x = vd_0to5, y = fit, lty = dts_cat), size = 2) +
geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit), colour = NA, alpha = 0.1) +
scale_color_manual(breaks = c("close to human settlement", "far from human settlement"), values = c("red", "blue")) + 
scale_fill_manual(breaks = c("close to human settlement", "far from human settlement"), values = c("red", "blue")) +
ylab("probability of succesful reproduction") + 
xlab("understory density") + 
theme_classic(15) +                  
theme(legend.position = c(0.8, 0.2), legend.title = element_blank(), legend.key.size = unit(2, 'lines'))
```

9. Figure 2:

```{r}
## Calculate cor values used for vertical lines below:
vert_lin <- r.all_rad$cor[r.all_rad$radius %in% c(78, 202, 450)]

ggplot(r.all_rad, aes(x = cor, y = R2m)) +
geom_point(size = 4) +
scale_x_reverse() +
geom_vline(xintercept = vert_lin, color = "black", linetype = "dashed") +
geom_text(aes(x = vert_lin[2] + 0.025, label = "450 m", y = 0.115), angle = 0, size = 5) +
geom_text(aes(x = vert_lin[3] + 0.021, label = "80 m", y = 0.115), angle = 0, size = 5) +
geom_text(aes(x = vert_lin[1] + 0.025, label = "200 m", y = 0.115), angle = 0, size = 5) +
scale_color_grey(start = 0.1, end = 0.5) + 
scale_fill_grey(start = 0.1, end = 0.5) +
xlab("correlation of ud at nest with ud at radius i around the nest") + 
ylab("r-squared with ud at radius i") + 
theme_classic(15)
```

