---
title: "Remotely sensed forest understory density and nest predator occurrence interact to predict suitable breeding habitat and the occurrence of a resident boreal bird species - Part 3"
author: "Julian Klein"
date: "14 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/RDefaultDirectory/RProjects/Lavskrika/lavskrika_rep_success")
```

# Part 3: Comparison between model predictions for breeding success and siberian jay occurrence.

The is based on rasters data extracted from .las files with FUSION. As raw LiDAR data cannot be redistributed part 3 cannot be run without first downloading LiDAR data from www.lantmateriet.se. The download is not free.

The raster showing Siberian jay occurrences will have to be required at Bradter et al.

The raster showing distance to settlements will have to be required at Statistics Sweden (SCB)

The analysis presented here is based on habitat data for 15 m around the nest.
The analysis for 80 m around the nest is exactly the same and correlated with 0.84

# Start:

1. Load all packages:

```{r, message=F, warning=F}
library(raster)
library(Metrics)
```

2. Define all functions:

```{r}
## Define the formula for predicting nest success
## Results from rep_succ_analysis with 15m rad
prob_success <- function(x, y) {
  
  p <- 0.1598 + 
    0.8465 * y + ## y = dts_far
    0.3624 * x - ## x = centered(log(vd_0to5))
    1.4957 * y * x
  
  return(exp(p)/(1+exp(p)))
  
}

## Define dts categorisation distance
R <- 1450

## Define focal rad
rad <- 15
```

3. Load all data:

```{r}
lp_shape <- shapefile("data/lp_shape.shp") ## Shape of the chosen area
## p.fit.data is the prediction data with the lme4 package produced in part 2 and reimported into part 3 here
p.fit.data <- read.csv("data/p.vd0t5_log_c.csv")
head(p.fit.data)
```

```{r}
vd_lp <- raster("data/Dens_05_5_mosaic.tif")*100
## *100 because input between 0 and 1 but formula expects between 0 and 100
plot(vd_lp); plot(lp_shape, add = TRUE)
```

```{r}
dts <- raster("data/dts.tif")
plot(dts); plot(lp_shape, add = TRUE)
```

```{r}
sj_occ <- raster("data/SFBestCellNbhMainland_99TM.tif")
plot(sj_occ); plot(lp_shape, add = TRUE)
```

4. Compare the predictions with the formula with the predictions from predict() to test if the formula is correct:

```{r}
## 
p.fit.data$dts_far[p.fit.data$dts_cat == "close"] <- 0
p.fit.data$dts_far[p.fit.data$dts_cat == "far"] <- 1

## C is needed in prob_success
C <- mean(log(p.fit.data$vd_0to5), na.rm = TRUE)

D_pred <- cbind("fit_formula" = prob_success(x = log(p.fit.data$vd_0to5) - C, y = p.fit.data$dts_far), p.fit.data)

plot(D_pred$fit, D_pred$fit_formula); abline(0, 1, col = "red")
```

5.  Adjust the raster files and predict breeding success for Arvidsjaur, Malå, Lycksele and Åsele

```{r}
## Reduce vd_lp to the extent of dts
vd_lp <- crop(vd_lp, extent(dts))

## Average vd_lp within radius used for the prediction in the study

## Define raster with value = 1 for all cells where understorey is not NA
num_vd_lp <- vd_lp 
num_vd_lp[!is.na(num_vd_lp)] <- 1 

#Define focal filter
f <- focalWeight(vd_lp, d = rad, type = "circle") 
f[f > 0] <- 1 

#Calculate sum of raster cells in focal filter
vd_lp_sum <- focal(vd_lp, f, fun = sum, na.rm = TRUE) 

#Calculate number of raster cells in focal filter
vd_lp_num <- focal(num_vd_lp, f, fun = sum, na.rm = TRUE)

## Calculate mean within rad
vd_lp_mean <- vd_lp_sum / vd_lp_num

## Calculate centered log of vd_lp_mean
vd_lp_log_cent <- log(vd_lp_mean)-C 

## Change the resolution of dts
dts <- disaggregate(dts, fact = 4)

## Categorise dts
dts_cat <- dts >= R
```

```{r, echo=FALSE}
invisible(gc())
```

```{r}
## Predict p(successful reprduction)
## Overlay fastest. Tested against calc and manual calculation. Results of all three the same
t1 <- Sys.time() 
lp_out <- overlay(vd_lp_log_cent, dts_cat, fun = prob_success)
t2 <- Sys.time()
t2-t1

## All NA's in lp_out become 0 as they are non-habitat and no reproduction is expected there
lp_out[is.na(lp_out)] <- 0
```

```{r, echo=FALSE}
invisible(gc())
```

```{r}
## Reduce lp_out to range found in study

range_far <- c(min(p.fit.data$vd_0to5[p.fit.data$dts_far == 1]), max(p.fit.data$vd_0to5[p.fit.data$dts_far == 1]))
range_close <- c(min(p.fit.data$vd_0to5[p.fit.data$dts_far == 0]), max(p.fit.data$vd_0to5[p.fit.data$dts_far == 0]))

## Define rasters that will become NA after the prediction
NA_far <- vd_lp_mean <= range_far[1] | vd_lp_mean >= range_far[2]
NA_close <- vd_lp_mean <= range_close[1] | vd_lp_mean >= range_close[2]
```

```{r}
## Set non study range predictions to 0. Not selected => no reproduction
lp_out[dts_cat == 1 & NA_far] <- 0
lp_out[dts_cat == 0 & NA_close] <- 0
```

5. Compare P(succ_repr) with p(Occurence): 

```{r, echo=FALSE}
invisible(gc())
```

```{r}
## Resample sj_occ to resolution of lp_out
lp_out_resamp <- resample(lp_out, sj_occ)

## Reduce both to the Kommun borders of Arvidsjaur, Åsele. Malå and Lycksele
lp_out_resamp <- mask(lp_out_resamp, lp_shape)
sj_occ <- mask(sj_occ, lp_shape)

## Compare lp_out with sj_occ:
cor.test(sj_occ[], lp_out_resamp[])

## Look at disagreement after normalisation of both rasters

lp_norm <- (lp_out_resamp - min(lp_out_resamp[], na.rm = TRUE))/(max(lp_out_resamp[], na.rm = TRUE) - min(lp_out_resamp[], na.rm = TRUE))
sjo_norm <- (sj_occ - min(sj_occ[], na.rm = TRUE))/(max(sj_occ[], na.rm = TRUE) - min(sj_occ[], na.rm = TRUE))

disag <- abs(sjo_norm - lp_norm)

plot(crop(disag, extent(lp_shape)), col = bpy.colors())
```

